---
title: "R Learner's Permit Exam"
author: "Andrew J. Greenlee, Ph.D."
date: "2/19/2021"
output:
  html_notebook:
    toc: yes
---

# Introduction
Over the course of the first few weeks of our class, you have learned some basic elements of the R programming language and RStudio IDE. As we move forward with our class, we will continue to add knowledge of new packages, tools, and data within R.

This R Learner's Permit is designed to assess your knowledge of the core elements of the R language and software that we have focused on so far in class.

## Instructions

Please follow the instructions contained in this R Notebook, which involve filling in code and generating output. Most instructions ask you to add or fill in code chunks. Others ask you to provide a written interpretation in the notebook portion of the document.

Complete as many items as you can. If you run into trouble completing an item, add comments to your code or in the notebook describing where you are running into problems, and what you think the next step might be to solving the problem you're having.

Complete as many items as you can and then upload your R Notebook to the "R Learners Permit" folder in your assignment repository on GitHub.

## The Data

In the "data" subfolder of this project, you will find one .rds file called NYC_Violations.rds. This is the data you will use to earn your learner's permit. These are "real world" data derived from two sources:

1. New York City [housing maintainance code violation](https://data.cityofnewyork.us/Housing-Development/Housing-Maintenance-Code-Violations/wvxf-dwi5) data for the years 2009 to 2019
2. Data on property characteristics from the [Primary Land Use Tax Lot Output (PLUTO)](https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-pluto-mappluto.page) database

I have merged selected columns from these datasets for you (you will find data documentation and metadata for both datasets in the documentation subfolder), and after taking a close look at the dataset and the documentation, you should be able to identify which portions of the data you will work with come from each dataset.

The PLUTO data contains information on each [tax lot](http://gis.nyc.gov/taxmap/map.htm) within the city of New York. By merging these data which contain characteristics of buildings and land with the building code violation data, the data you have to work with describes the characteristics of the building and property where the violation occurred as well as the nature of the violation. It is essential for you to spend the time to think carefully about what each row in the building code violation data represents.

# Earn Your Learner's Permit

With that background in mind, its time to earn your learner's permit.

## Task 1: Load and Describe the Data

At the beginning of the script, we will create a holding area within which you can load any packages you may need to work with as you year your R Learner's Permit (if you are not using any packages, please leave this area blank):
```{r, echo=FALSE}
# Your Code Here
```

Use the `readRDS()` command to load the data into an object with the name of your choice.
```{r}
# Your Code Here
NYC_vio = readRDS("data/NYC_Violations_WB.rds")
```

Describe the structure of the data, including the number of rows and columns in the data. Create a table outside of the code chunk below in your notebook that describes each variable name and what each row in the data represents (a [Markdown table generator](https://www.tablesgenerator.com/markdown_tables) may be useful):
```{r}
# Your Code Here
nrow(NYC_vio)
ncol(NYC_vio)
```

**Your Narrative Description of the Data Structure:**
| BBL                                    | Borough         | NTA                              | ViolationStatus                 | InspectionyYear    | unitstital                | ownername | assessland | assesstot | yearbuilt           | histdist         | ordernumber             |
|----------------------------------------|-----------------|----------------------------------|---------------------------------|--------------------|---------------------------|-----------|------------|-----------|---------------------|------------------|-------------------------|
| unique identifier of borough,block,lot | name of borough | the neiborghbood tabulation area | status of violation(open/close) | year of inspection | the total number of units | ownername |            |           | built in which year | history district | refer to a specific law |

***
## Task 2: Summarize Violations

Describe the number of unique violations for each year (information on the year the unit was inspected for violations is contained in the InspectionYear variable):
```{r}
# Your Code Here
library(tidyverse)
```

```{r}
summary_year <- NYC_vio %>%
  group_by(InspectionYear) %>%
  summarise(yearly_tot = sum(unitstotal,na.rm = TRUE))
```

Now, build upon your summary code to make a plot (geometry of your choice) of inspections by year:
```{r}
library(ggplot2)
ggplot(data=summary_year, aes(x=InspectionYear,y=yearly_tot))+
  geom_bar(stat="identity")
```

In your own words, what is the overall trend with regards to the number of violations issued between 2009 and 2019?:

**Your Narrative Description:** 
the number of vialation is growing more and more rapidly.

Next, describe the number of unique violations for each New York City borough. Do this using both *base* r code and using `dplyr` code that utilizes the `summarise()` function:

#use base R ???

```{r}
# Your Code Here
sum_tot = rep(0,5)
for(i in 1:5){
  sum_tot[i] = sum(NYC_vio$unitstotal[NYC_vio$Borough == unique(NYC_vio$Borough)[i]], na.rm = TRUE)
}
names(sum_tot) = unique(NYC_vio$Borough)
```

#using `dplyr` code
```{r}
summary_borough <- NYC_vio %>%
  group_by(Borough) %>%
  summarise(tot_vio = sum(unitstotal,na.rm = TRUE))
```

***
## Task 3: More Detail on Violations

Let's create some more detailed summaries of violations. 

Create a table that enumerates by borough 
the number of violations 
that are currently closed or open (ViolationStatus). 
Remove the `NA` values so that you are just reporting open and closed cases:

```{r}
# Your Code Here
B_status <- NYC_vio %>%
  group_by(Borough) %>%
  summarise(B_closed = sum(unitstotal[ViolationStatus == "Close"], na.rm = TRUE),
            B_open = sum(unitstotal[ViolationStatus == "Open"], na.rm = TRUE))
```

See if you can refine the formatting of this table. Although we haven't formally introduced `pivot_wider()` yet, try using it so that you produce a summary 
in which there are three columns - the borough, the number of closed cases, and the number of open cases (remember you can search for documentation, 

e.g. `?pivot_wider()`). You will need to supply pivot_wider with an argument for `names_from` and an argument for `values_from`. Names should come from the Violation Status column, and values should come from the column in which the count of violations is contained.
#pivot_wider()
```{r}
# Your Code Here
B_status2 <- B_status %>%
  pivot_wider(names_from = "Borough", 
              values_from ="B_open")
```
Now create a chart (geom of your choice) that represents 
the number of *open cases* 
by borough 
by the year in which the violation occurred:
#every year, the total number of的open cases in each borough，(x=year,y=case number)
```{r, message=FALSE}
BYOPEN <- NYC_vio %>%
  filter(ViolationStatus == "Open") %>%
  group_by(Borough,InspectionYear) %>%
  summarise(BY_open = sum(unitstotal,na.rm = TRUE))
```

```{r}
ggplot(data = BYOPEN, aes(x=InspectionYear,y=BY_open,color=Borough)) +
         geom_line()
```

Now, create a table with the total number of *buildings* with violations (remember, there may be multiple violations per building in the data). Remember that the BBL column contains a unique identification number for each building (technically it's each tax lot, but we'll use that as a proxy for building). You can use the `distinct()` command to create a subset of the distinct buildings with violations. You could also use the `summarise()` command to accomplish something similar...

```{r}
# Your Code Here
NYC_vio %>%
  filter(ViolationStatus=="Open") %>%
  distinct(BBL, .keep_all = TRUE)
  
```

Let's create a final summary table of buildings with violations and total violations by borough. Use `dplyr` [join syntax](https://dplyr.tidyverse.org/reference/join.html) to accomplish this. 

Your table should have a column containing the borough name, 
a column containing the total number of buildings with violations, 
and a column containing the total number of violations.
```{r}
# Your Work Here
Tab.1 <- NYC_vio %>%
  filter(ViolationStatus=="Open") %>%
  distinct(BBL, .keep_all = TRUE) %>%
  group_by(Borough) %>%
  summarise(tot_num_vio=length(BBL))

B_vio <- inner_join(Tab.1, summary_borough, by = "Borough")
```

Finally, add to this table, the average violations per building by borough. You can use the `mutate()` command to add this column to your existing table.

```{r}
# Your Work Here
B_vio %>%
  mutate(avg_vio=tot_vio/tot_num_vio)
```
Which boroughs contain higher rates of violations per building?
BRONX
**Your Narrative Description on Violations by Borough:**
BRONX has the highest rate of violations, QUEENS has the lowest rate.
***

## Task 4: Building Ages

Create a table that displays the mean and median ages (yearbuilt) of buildings with violations by borough. Remember that individual buildings may have multiple observations - you will need to figure out how to create a summary of the dataset that includes one record per building to generate an unbiased summary.
```{r}
# Your Code Here
NYC_vio %>%
  filter(yearbuilt>0) %>%
  distinct(BBL, .keep_all = TRUE) %>%
  group_by(Borough) %>%
  summarise(mean_age=mean(yearbuilt, na.rm=TRUE),
            med_age=median(yearbuilt, na.rm=TRUE),
            min_age=min(yearbuilt,na.rm=TRUE),
            max_age=max(yearbuilt,na.rm=TRUE))
```

Add to the table that you just made two columns that include the ages of the oldest and youngest buildings for each borough (since the years are numeric, the functions `min()` and `max()` may be useful here):
```{r}
# Your Code Here

```

What do you discover when trying to do this, and how might you modify your code to rectify it?

**Your Narrative Description of this Table and code to address your discovery:**
this column use 0 ro represent missing value, use filter()
```{r}
# Your Code here
NYC_vio %>%
  filter(yearbuilt>0) %>%
  distinct(BBL, .keep_all = TRUE) %>%
  group_by(Borough) %>%
  summarise(mean_age=mean(yearbuilt, na.rm=TRUE),
            med_age=median(yearbuilt, na.rm=TRUE),
            min_age=min(yearbuilt,na.rm=TRUE),
            max_age=max(yearbuilt,na.rm=TRUE))
```

**Your Narrative Description of the implications of any modifications you made between the last two code chunks:**
use filter()

***
## Task 5: Decade Comparison
Thinking about each borough, create a table that would allow us to compare the frequency of violations in 2009 with the frequency of violations in 2019. 

```{r}
# Your Code Here
NYC_vio %>%
  filter((InspectionYear==2009)|(InspectionYear==2019))%>%
  group_by(Borough) %>%
  summarise(vio_2009=sum(unitstotal[InspectionYear==2009],na.rm = TRUE),
            vio_2019=sum(unitstotal[InspectionYear==2019],na.rm = TRUE))
```

How has the frequency of violations changed over the decade?
**Your Narrative Description:**

***

## Task 6: Frequency of Violations by Violation Type

The OrderNumber column contains information on the actual violation recorded (the housing maintenance code violation user guide contains a narrative description of each order number).

What is the most frequent code violation recorded in each borough?
```{r}
# Your Code Here
install.packages("modeest")
```
```{r}
library(modeest)
```
```{r}
NYC_vio %>%
  group_by(Borough) %>%
  summarise(mfv(OrderNumber))
```

**Your Narrative Description:**
divide the data accoding to the name of borough, then try to find the most request code in the column named OrderNumber.
***
## Task 7: Building Owners with the Most Violations

Which identifiable owner (ownername) has the highest rate of violations per unit?
#group the data according to ownername, sum the total number of the nuit for each person and the total number of violations, count the rate, and then max to find the largest number.
```{r}
# Your Code Here

```

**Your Narrative Description:**

***
## Task 8: Buildings in Historic Districts

For each Borough, how many violations occur in *buildings* built after 1957 which are located in historic districts (histdist)?

```{r}
# Your Code Here
NYC_vio1 <- NYC_vio %>%
  mutate( histdist= ifelse(is.na(histdist),0,1))
```

```{r}
B_his <- NYC_vio1 %>%
  filter(yearbuilt>1957) %>%
  group_by(Borough) %>%
  summarise(his_vio_tot=sum(unitstotal)[histdist=1], na.rm = TRUE)
```

Add to this table the average assessed value of the land (assessland), the average assessed value of the buildings, and the average total assessed value.

```{r}
# Your Code Here
ave_land_value <-  NYC_vio1 %>%
  filter(yearbuilt>1957) %>%
  group_by(Borough) %>%
  summarise(ave_land= mean(assessland[histdist=1]))
```

```{r}
ave_b_value <-  NYC_vio1 %>%
  filter(yearbuilt>1957) %>%
  distinct(BBL, .keep_all = TRUE) %>%
  group_by(Borough) %>%
  summarise(ave_land= mean(assesstot[histdist=1]))
```

```{r}
ave_value <- inner_join(ave_land_value, ave_b_value, by = "Borough")
```

For those buildings in historic districts, how different are the ratios of the assessed values for land and buildings in each borough? What is your interpretation of these differences?

```{r}
# Your Code Here
ave_value %>%
  mutate(ratio=ave_land.x/ave_land.y)
```
**Your Narrative Description:**
compare with the value of land in other area, the value of land in manhattan is extremely high.
But i don't know how this ratio can tell the story about the value of land and building.
***
# The End
That's it - you're done! Your permit will be assessed and returned shortly. Please submit your R notebook to the appropriate GitHub folder. Rename the notebook so that the name is lastname_firstname_Permit.Rmd (e.g. Greenlee_Andrew_Permit.Rmd). You just need to upload the .Rmd file - no need to include the whole file structure.